---
title: "Assingment3_Zhou_HealthyRicePrediction"
author: "Yan Zhou"
date: "11/18/2021"
output: html_document
```{r}
library(keras)
library(tidyverse)

```
# to store a sebset of data that we are going to use
```{r}
base_dir <- "rice_images/health_unhealth_total" # to store a sebset of data that we are going to use
dir.create(base_dir)


train_dir <- file.path(base_dir, "train")
dir.create(train_dir)

validation_dir <- file.path(base_dir, "validation")
dir.create(validation_dir)

test_dir <- file.path(base_dir, "test")
dir.create(test_dir)
```
# create folders saving healthy images and unhealthy images
```{r}
train_health_dir <- file.path(train_dir, "health")
dir.create(train_health_dir)
train_unhealth_dir <- file.path(train_dir, "unhealth")
dir.create(train_unhealth_dir)
validation_health_dir <- file.path(validation_dir, "health")
dir.create(validation_health_dir)
validation_unhealth_dir <- file.path(validation_dir, "unhealth")
dir.create(validation_unhealth_dir)
test_health_dir <- file.path(test_dir, "health")
dir.create(test_health_dir)
test_unhealth_dir <- file.path(test_dir, "unhealth")
dir.create(test_unhealth_dir)
```

# split healthy images 70% for training, 15% for validation and 15% for testing.
```{r}

num_health<-length(list.files("rice_images/_Healthy"))
num_health_train <- ceiling(num_health*0.7)
num_health_train
num_health_validation <- ceiling(num_health*0.85)
num_health_validation
health_dataset_dir1 <- "rice_images/_Healthy"
fnames <- paste0("shape ", 1:num_health_train, " .jpg")
file.copy(file.path(health_dataset_dir1, fnames), file.path(train_health_dir))
fnames <- paste0("shape ", (num_health_train + 1):num_health_validation, " .jpg")
file.copy(file.path(health_dataset_dir1, fnames), file.path(validation_health_dir))
fnames <- paste0("shape ", (num_health_validation + 1):num_health, " .jpg")
file.copy(file.path(health_dataset_dir1, fnames), file.path(test_health_dir))

```

# copy original folder "_BrownSpot" to "new_BrownSpot" and rename the files in the folder "new_BrownSpot" 


```{r}
base_dir <- "rice_images/new_BrownSpot" # to store a sebset of data that we are going to use
dir.create(base_dir)
num_brown<-length(list.files("rice_images/_BrownSpot"))
brown_dataset_dir<-"rice_images/_BrownSpot"
fnames <- paste0("shape ", 1:num_brown, " .jpg")
file.copy(file.path(brown_dataset_dir, fnames), file.path(base_dir))


my_path <- "rice_images/new_BrownSpot/"
file_names_old <- list.files(my_path)              # Get current file names
file_names_new <- paste0("shape_brown ",               # Define new file names
                         1:length(file_names_old),
                         ".jpg")
file.rename(paste0(my_path, file_names_old),       # Rename files
            paste0(my_path, file_names_new))


```
# copy original folder "_Hispa" to "new_Hispa" and rename the files in the folder "new_Hispa"

```{r}
base_dir <- "rice_images/new_Hispa" # to store a sebset of data that we are going to use
dir.create(base_dir)
num_hispa<-length(list.files("rice_images/_Hispa"))
hispa_dataset_dir<-"rice_images/_Hispa"
fnames <- paste0("shape ", 1:num_hispa, " .jpg")
file.copy(file.path(hispa_dataset_dir, fnames), file.path(base_dir))



my_path <- "rice_images/new_Hispa/"
file_names_old <- list.files(my_path)              # Get current file names
file_names_new <- paste0("shape_hispa ",               # Define new file names
                         1:length(file_names_old),
                         ".jpg")
file.rename(paste0(my_path, file_names_old),       # Rename files
            paste0(my_path, file_names_new))
            
```

# copy original folder "_LeafBlast" to "new__LeafBlast" and rename the files in the folder "new_LeafBlast"

```{r}
base_dir <- "rice_images/new_LeafBlast" # to store a sebset of data that we are going to use
dir.create(base_dir)
num_Leafblast<-length(list.files("rice_images/_LeafBlast"))
leafblast_dataset_dir<-"rice_images/_LeafBlast"
fnames <- paste0("shape ", 1:num_Leafblast, " .jpg")
file.copy(file.path(leafblast_dataset_dir, fnames), file.path(base_dir))


my_path <- "rice_images/new_LeafBlast/"
file_names_old <- list.files(my_path)              # Get current file names
file_names_new <- paste0("shape_leafblast ",               # Define new file names
                         1:length(file_names_old),
                         ".jpg")
file.rename(paste0(my_path, file_names_old),       # Rename files
            paste0(my_path, file_names_new))
```
Brownspot images: 70% for training, 15%for validation and 15% for testing
Hispa images: 70% for training, 15%for validation and 15% for testing
Leafblast images: 70% for training, 15%for validation and 15% for testing
take all training data of Brownspot images, Hispa images,Leafblast images as unhealthy training data
take all validation data of Brownspot images, Hispa images,Leafblast images as unhealthy validation data
take all test data of Brownspot images, Hispa images,Leafblast images as unhealthy test data

add files of new_LeafBlast into unhealthy folders
```{r}
unhealth_dataset_dir1 <- "rice_images/new_LeafBlast"

num_unhealth<-length(list.files(unhealth_dataset_dir1))
num_unhealth_train <- ceiling(num_unhealth*0.7)
num_unhealth_train
num_unhealth_validation <- ceiling(num_unhealth*0.85)
num_unhealth_validation


fnames <- paste0("shape_leafblast ", 1:num_unhealth_train, ".jpg")
file.copy(file.path(unhealth_dataset_dir1, fnames), file.path(train_unhealth_dir))
fnames <- paste0("shape_leafblast ", (num_unhealth_train + 1):num_unhealth_validation, ".jpg")
file.copy(file.path(unhealth_dataset_dir1, fnames), file.path(validation_unhealth_dir))
fnames <- paste0("shape_leafblast ", (num_unhealth_validation + 1):num_unhealth, ".jpg")
file.copy(file.path(unhealth_dataset_dir1, fnames), file.path(test_unhealth_dir))

```
#add files of  _BrownSpot into unhealthy folders

```{r}
unhealth_dataset_dir1 <- "rice_images/new_BrownSpot"
num_unhealth<-length(list.files(unhealth_dataset_dir1))
num_unhealth_train <- ceiling(num_unhealth*0.7)
num_unhealth_train
num_unhealth_validation <- ceiling(num_unhealth*0.85)
num_unhealth_validation


fnames <- paste0("shape_brown ", 1:num_unhealth_train, ".jpg")
file.copy(file.path(unhealth_dataset_dir1, fnames), file.path(train_unhealth_dir))
fnames <- paste0("shape_brown ", (num_unhealth_train + 1):num_unhealth_validation, ".jpg")
file.copy(file.path(unhealth_dataset_dir1, fnames), file.path(validation_unhealth_dir))
fnames <- paste0("shape_brown ", (num_unhealth_validation + 1):num_unhealth, ".jpg")
file.copy(file.path(unhealth_dataset_dir1, fnames), file.path(test_unhealth_dir))


```
#add files of  _Hispa into unhealthy folders
```{r}
unhealth_dataset_dir1 <- "rice_images/new_Hispa"
num_unhealth<-length(list.files(unhealth_dataset_dir1))
num_unhealth_train <- ceiling(num_unhealth*0.7)
num_unhealth_train
num_unhealth_validation <- ceiling(num_unhealth*0.85)
num_unhealth_validation


fnames <- paste0("shape_hispa ", 1:num_unhealth_train, ".jpg")
file.copy(file.path(unhealth_dataset_dir1, fnames), file.path(train_unhealth_dir))
fnames <- paste0("shape_hispa ", (num_unhealth_train + 1):num_unhealth_validation, ".jpg")
file.copy(file.path(unhealth_dataset_dir1, fnames), file.path(validation_unhealth_dir))
fnames <- paste0("shape_hispa ", (num_unhealth_validation + 1):num_unhealth, ".jpg")
file.copy(file.path(unhealth_dataset_dir1, fnames), file.path(test_unhealth_dir))

```

# count how many pictures are in each split folder

```{r}
cat("total training health images:", length(list.files(train_health_dir)), "\n")
cat("total validation health images:", length(list.files(validation_health_dir)), "\n")
cat("total test health images:", length(list.files(test_health_dir)), "\n")


cat("total training unhealth images:", length(list.files(train_unhealth_dir)), "\n")
cat("total validation unhealth images:", length(list.files(validation_unhealth_dir)), "\n")
cat("total test unhealth images:", length(list.files(test_unhealth_dir)), "\n")

```            
#Build CNN model
```{r}
library(keras)
model_v1 <- keras_model_sequential() %>%
  layer_conv_2d(filters = 32, kernel_size = c(3, 3), activation = "relu",
                input_shape = c(150, 150, 3)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_flatten() %>%
  layer_dense(units = 512, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")


model_v1 %>% compile(
  loss = "binary_crossentropy",
  optimizer = optimizer_rmsprop(),
  metrics = c("acc")
)

summary(model_v1)
```
#Data preprocessing
```{r}
train_datagen <- image_data_generator(rescale = 1/255)
validation_datagen <- image_data_generator(rescale = 1/255)
train_generator <- flow_images_from_directory(
  train_dir, # Target directory
  train_datagen, # Training data generator
  target_size = c(150, 150), # Resizes all images to 150 × 150
  batch_size = 20, # 20 samples in one batch
  class_mode = "binary" # Because we use binary_crossentropy loss,
  # we need binary labels.
)
validation_generator <- flow_images_from_directory(
  validation_dir,
  validation_datagen,
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = "binary"
)
```

#fit the model
```{r, eval=FALSE}
history_v1 <- model_v1 %>%
  fit_generator(
    train_generator,
    steps_per_epoch = 50,
    epochs = 20,
    validation_data =
      validation_generator,
    validation_steps = 30
  )


model_v1 %>% save_model_hdf5("Assignment3-Zhou-RicePrediction-model-v1.h5")
write_rds(history_v1, "Assignment3-Zhou-RicePrediction-history-v1.rds")
plot(history_v1)
```

```{r}

history_v1 <- read_rds("Assignment3-Zhou-RicePrediction-history-v1.rds")
plot(history_v1)

```
#evaluate this model using the test data

```{r}
test_datagen <- image_data_generator(rescale = 1/255)

test_generator <- flow_images_from_directory(
  test_dir,
  test_datagen, 
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = "binary"
)


model_v1 <- load_model_hdf5("Assignment3-Zhou-RicePrediction-model-v1.h5")

model_v1 %>% evaluate_generator(test_generator, steps = 50)

```
#Setting up a data augmentation configuration via image_data_generator
```{r}
datagen <- image_data_generator(
  rescale = 1/255,
  rotation_range = 40, # randomly rotate images up to 40 degrees
  width_shift_range = 0.2, # randomly shift 20% pictures horizontally
  height_shift_range = 0.2, # randomly shift 20% pictures vertically
  shear_range = 0.2, # randomly apply shearing transformations
  zoom_range = 0.2, # randomly zooming inside pictures
  horizontal_flip = TRUE, # randomly flipping half the images horizontally
  fill_mode = "nearest" # used for filling in newly created pixels
)
```
#Defining a new convnet that includes dropout

```{r}
model_v2 <- keras_model_sequential() %>%
  layer_conv_2d(filters = 32, kernel_size = c(3, 3), activation = "relu",
                input_shape = c(150, 150, 3)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_flatten() %>%
  layer_dropout(rate = 0.5) %>% # randomly set 50% of weights to 0
  layer_dense(units = 512, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")
model_v2 %>% compile(
  loss = "binary_crossentropy",
  optimizer = optimizer_rmsprop(lr = 1e-4),
  metrics = c("acc")
)
```
#Training the convnet using data-augmentation generators:
```{r}
test_datagen <- image_data_generator(rescale = 1/255) # no data augmentation
train_generator <- flow_images_from_directory(
  train_dir,
  datagen, # Our data augmentation configuration defined earlier
  target_size = c(150, 150),
  batch_size = 32,
  class_mode = "binary"
)
validation_generator <- flow_images_from_directory(
  validation_dir,
  test_datagen, # Note that the validation data shouldn’t be augmented!
  target_size = c(150, 150),
  batch_size = 32,
  class_mode = "binary"
)

```
#fit the model
```{r, eval=FALSE}
history_v2 <- model_v2 %>% fit_generator(
  train_generator,
  steps_per_epoch = 100,
  epochs = 50,
  validation_data = validation_generator,
  validation_steps = 50
)


model_v2 %>% save_model_hdf5("Assignment3-Zhou-RicePrediction-model-v2.h5")
write_rds(history_v2, "Assignment3-Zhou-RicePrediction-history-v2.rds")
plot(history_v2)

```

```{r}
history_v2 <- read_rds("Assignment3-Zhou-RicePrediction-history-v2.rds")

plot(history_v2)

```

#evaluate this model using the test data

```{r}
test_generator <- flow_images_from_directory(
  test_dir,
  test_datagen, 
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = "binary"
)


model_v2 <- load_model_hdf5("Assignment3-Zhou-RicePrediction-model-v2.h5")

model_v2 %>% evaluate_generator(test_generator, steps = 50)
            
```
                        
                                    
                                                
                                                            
                                                                        
                                                                                    
                                                                                                
                                                                                                            
                                                                                                                        
                                                                                                                                    
                                                                                                                                                
                                                                                                                                                                        
            
            